---
title: "AtlantisOMtoSSVignette"
author: "Christine Stawitz, Sarah Gaichas"
date: "May 8, 2019"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: true
bibliography: refs.bib
csl: cjfas.csl
vignette: >
  %\VignetteIndexEntry{AtlantisOMtoSSVignette}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

First, you will want to load packages and initialize the correct file and working directory names.

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(data.table)
library(here)
```

If you want to load the local of the atlantisom R package, use devtools:

```{r message=FALSE, warning=FALSE}
#library(devtools)
#package.dir <- "C:/Users/chris/Documents/Github/atlantisom"
#package.dir <- here()
#devtools::load_all(package.dir)
library(atlantisom)
```

Or you can install directly from the Github repository.

```{r eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("r4atlantis\atlantisom")
```

## Initializing input files and directories

You will first need to tell `atlantisom` to know where to look for the output and input files from your atlantis model run. Here, we will give the directory where the Atlantis inputs and outputs are stored `d.name`, the location of the functional groups file `functional.group.file` (.csv), the biomass pools file `biomass.pools.file` (.nc), the box locations file `box.file` (.bgm), an `initial.conditions.file` (.nc), the biology .prm file `biol.prm.file` (.prm), and the run .prm file `run.prm.file` (.prm). You will also need to specify a scenario name, which will be used to define the output files (i.e. output is stored in a number of netCDF files of the format: output<scenario><value>.nc). All of these files should be stored in `d.name`.

```{r initialize, eval = FALSE}
#ss.directory <- here("inst/extdata/Sardine_SS_files")

#Need a way to manage data that doesn't rely on specific paths
d.name <- here("non_synced_data/CalCurrentSummitScenario1")
source(here("config/CCConfig.R"))
```

## Getting the "true" operating model values

There are a number of functions in the package that begin with the prefix `load` that load various files. See documentation if you'd only like to load one file. The `atlantisom::run_truth()` function uses the above file definitions and calls a number of the `load` functions to read in all of the atlantis output. Note: this call reads in a number of large .nc files, so it will take a few minutes to return.

```{r get_truth, , eval = FALSE, message=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

if(!file.exists(file.path(d.name,"outputCCV3run_truth.RData"))){
#Store all loaded results into an R object
results <- run_truth(scenario = scenario.name,
          dir = d.name,
          file_fgs = functional.groups.file,
          file_bgm = box.file,
          select_groups = funct.group.names,
          file_init = initial.conditions.file,
          file_biolprm = biol.prm.file,
          file_runprm = run.prm.file
          save(results, file = "outputCCV3run_truth.RData")
)
} else{
  results <- get(load(file.path(d.name,"outputCCV3run_truth.RData")))
}

```

Now the R object `results` with the comprehensive results from the Atlantis model has been read in. 

## Simulating "sampling" from an Atlantis OM

There are a number of functions to "sample" biological data from this comprehensive Atlantis output. These include `create_survey`, `sample_ages`, `sample_diet`, `sample_fish`, and `sample_survey_biomass`. Below, we use the `sample_survey_biomass` function to extract time series of a biomass index for a survey and fisheries CPUE for Pacific sardine. 

```{r, eval = FALSE, echo = TRUE}
species=c("Pacific_sardine")

####### Boxes = spatial area, use all for now
# effic - ?
# sel - age-based selectivity curve
boxes <- unique(results$nums$polygon)
effic <- data.frame(species=species, efficiency=0.5)
sel<- data.frame(species=rep(species,10),
               agecl=c(1:10),
              selex=rep(1,10))

#Extract survey numbers
tmp <- create_survey(dat=results$nums, time=timeall, species=species, boxes=boxes, effic=effic, selex=sel)

#wtAtAge comes from empirical data from assessment, in kg
wtAtAge <- data.frame(species=rep("Pacific_sardine",10),
                agecl=1:10,
                wtAtAge=c( 0.0542, 0.0837,	0.1103,	0.1323,	0.1497,	0.163,	0.1729,	0.1801,	0.1854,	0.1941))

#Set survey CV at 0.2
cv <- data.frame(species="Pacific_sardine", cv=c(0.2))

#Sample the survey biomass
survObsBiom <- sample_survey_biomass(dat=tmp,cv=cv,wtAtAge)

#Define selectivity curve for the fishery
selCatch<- data.frame(species=rep(species,10),
                 agecl=c(1:10),
                 selex=(1/(1+exp(-0.5*(1:10-2)))))

#How do we distinguish survey sampling from fisheries sampling? Maybe define by historical sampling locations?
fishery <- create_survey(dat=results$nums, time=seq(1,99,1), species=species, boxes=boxes, effic=effic, selex=selCatch)

# Make CV of fishery CPUE slightly larger
cv <- data.frame(species="Pacific_sardine", cv=c(0.4))

# Get catch biomass
catchObsBiom <- sample_survey_biomass(dat=fishery,cv=cv,wtAtAge)
```

We now have the full time series of catch and survey biomass from the atlantis output for Pacific sardine. Next, we read in the dummy SS data (`.dat`) file and replace the abundance indices with the sampled time series.

```{r, eval = FALSE, echo = TRUE}
#Read in dummy data
file <- system.file("extdata","Sardine_SS_files",file="sardEM_3_3.dat",package="atlantisom")
sardine.dat <- r4ss::SS_readdat_3.30(file)

#Replace CPUE observations for first time series with output from Atlantis - last 25 years
# SS units are metric tons, so we divide by 1000
sardine.dat$CPUE[1:75,"obs"] <- survObsBiom$atoutput[25:99]/1000

#Replace CPUE observations for second time series with output from Atlantis - last 25 years
# SS units are metric tons, so we divide by 1000
sardine.dat$CPUE[76:150,"obs"] <- catchObsBiom$atoutput[25:99]/1000
r4ss::SS_writedat_3.30(datlist = sardine.dat,
                       outfile = file.path(assessment_dir, "sardEM_3_3.dat"))
```

