---
title: "An introduction to atlantisom for simulating multispecies observations"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: true
bibliography: refs.bib
csl: cjfas.csl
vignette: >
  %\VignetteIndexEntry{An introduction to atlantisom}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Installing the atlantisom R package

atlantisom requires R version 3.1.0 or greater.
The package can be run on OS X, Windows, or Linux.
The development version of atlantisom can be installed from GitHub
with the following code:

```{r, install-and-load, eval=FALSE}
# install.packages("devtools") # if needed
devtools::install_github("r4atlantis/atlantisom", dependencies = TRUE)

# If you would like the vignettes available with the GitHub development version:
devtools::install_github("r4atlantis/atlantisom", dependencies = TRUE, build_vignettes = TRUE)
```

You can then load the package with:

```{r, load-package, eval=FALSE}
library("atlantisom")
```

You can read the help files and access the various vignettes (including this one) with:

```{r, help, eval=FALSE}
?atlantisom
help(package = "atlantisom")
browseVignettes("atlantisom")
```

See @fulton2010 for a more formal (but brief) overview of
[atlantis](http://atlantis.cmar.csiro.au/).

Please direct any questions, suggestions, or bug reports to the atlantisom
issue tracker: <https://github.com/r4atlantis/atlantisom/issues>, or to an
author as listed on the title page of this vignette.

If you use atlantisom in a publication, please cite the package as
indicated by running `citation("atlantisom")` in the R console:

```{r, citation, cache=FALSE}
citation("atlantisom")
```

# An overview of the atlantisom simulation structure

atlantisom is an R package to make it relatively quick and easy
to run simulations with the ATLANTIS output.

# References
