---
title: "atlantisom demo notebook"
output: 
  html_notebook: default
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
devtools::install_github("r4atlantis/atlantisom")
library(atlantisom)
```

## Introduction

This demonstrates the Atlantis to `atlantisom` to SS3 workflow with the California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_22". 

We will go from Atlantis model output to SS input and run in this page for a single species, "sardine", that doesn't require any age class splitting. 


```{r initialize}

species_ss <- c("Pacific_sardine")
age_classes <- 1:10

source(here("config/NEMOWConfig.R"))
needed.files <- ls()
print(unlist(lapply(needed.files, FUN=get)))

```

The below example requires a folder in your project directory named "atlantisom" containing the above files from an Atlantis output.

```{r get_names, message=FALSE, warning=FALSE}
#Check for needed files


#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

This can be run with `species_ss` in `select_groups` to get results only for that species. WARNING: running with all species takes a long time! (50 minutes for this output)

```{r get_truth_ss}
  #Store all loaded results into an R object
sardine_truth <- if(!file.exists(file.path(d.name, 
                          "sardine_truth.RData"))){
  #Store all loaded results into an R object
  sardine_truth <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = "Pacific_sardine",
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
  )
} else {
  sardine_truth <- get(load(file.path(d.name,
                              "sardine_truth.RData")))
}

```

This truth object shows the true population dynamics for sardine from Atlantis which we need for reference to compare with estimated population dynamics from the stock assessment estimation models. 

Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions. Here the survey functions are still separate.


## Derive “Data” from Atlantis to give to SS3

Specify our survey sampling. This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc. 

Here are specifications which approximate the true sardine acoustic survey

```{r sardine-survey-spec}

# generalized timesteps all models
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
midptyr <- round(median(seq(0,stepperyr)))

# a survey that takes place once per year mid year
annualmidyear <- seq(midptyr, noutsteps, stepperyr)
# learned the hard way this can be different from ecosystem outputs
fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))


source(here("config/sardine_survey.R"))

```

Get input survey biomass for SS (these have q, selectivity, cv)

```{r toSS-surveyts-comps}

#Sample survey - 3rd timestep simulates a summer survey
survey_out <- create_survey(dat=sardine_truth$nums, 
                            time=survey_sample_full, 
                            species=species, 
                            boxes=boxall, 
                            effic=effic, 
                            selex=sel)

#Try a biomass based survey for comparison
survey_outB <- create_survey(dat=sardine_truth$biomass_ages,
                            time=survey_sample_full,
                            species=species,
                            boxes=boxall,
                            effic=effic,
                            selex=sel)


#Set effective sample size for age compositions
effN <- surveyEffN
highEffN <- data.frame(species=species, effN=rep(effN, length(species)))

#Sample fish for age composition
age_comp_data <- sample_fish(survey_out, highEffN)

# aggregate true resn per survey design
survey_aggresnstd <- aggregateDensityData(dat = sardine_truth$resn,
                                          time = survey_sample_full,
                                          species = species,
                                          boxes = boxall)

# aggregate true structn per survey design
survey_aggstructnstd <- aggregateDensityData(dat =sardine_truth$structn,
                                             time = survey_sample_full,
                                             species = species,
                                             boxes = boxall)

ss_structnstd <- sample_fish(survey_aggstructnstd,
                             effN,
                             sample=FALSE)
ss_resnstd <- sample_fish(survey_aggresnstd,
                          effN,
                          sample=FALSE)

#Extract length composition data
ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
                                     resn = ss_resnstd,
                                     nums = age_comp_data,
                                     biolprm = sardine_truth$biolprm, fgs = sardine_truth$fgs,
                                     CVlenage = CVs$lenage, remove.zeroes=TRUE)

#Need to replace with interp function
wtAtAge <- ss_length_stdsurv$muweight %>%
  select(species, agecl, time, wtAtAge = atoutput) %>%
  mutate(wtAtAge = wtAtAge/1000)

# CV function
cv <- data.frame(species=species_ss, cv=CVs$survey)

#Sample survey biomass
survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)

# check against survey with truth$biomass_ages output
wtage1 <- data.frame(species=rep(species_ss, each=max(age_classes)),
                    agecl=rep(c(age_classes),length(species_ss)),
                    wtAtAge=rep(1000.0,length(species_ss)*max(age_classes)))

survObsBiomB <- sample_survey_biomass(dat=survey_outB,cv=cv,wtage1)

# survey numbers, not sure which SS needs?
survObsNum <- sample_survey_numbers(dat=survey_out,cv=cv)


```


Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age).

Because catch composition information goes in to the assessment as a proportion, we can use fishery catch at age from this legacy codebase even with absolute catch numbers likely half what they should be overall.


```{r toSS-fisherycomps}

# Survey length comps and wtage done above to get survey ts using nums*wtage approach

#We end up using CAAL for the survey below, so let's generate fishery age comps instead
effN <- fisheryEffN/fstepperyr
effN <- data.frame(species=species, effN=effN)

#catch at age each timestep summed over polygons
# catch at age by area and timestep
catch_numbers <-  create_fishery_subset(dat = sardine_truth$catch,
                                         time = fish_times,
                                         species = species,
                                         boxes = boxall)

catch_numsss_samp <- sample_fish(catch_numbers, effN)

rm(catch_numbers)
gc()
#Get weights
# aggregate true resn per fishery subset design
catch_aggresnss <- aggregateDensityData(dat = sardine_truth$resn,
                                 time = fish_times,
                                 species = species,
                                 boxes = boxall)

# aggregate true structn fishery subsetdesign
catch_aggstructnss <- aggregateDensityData(dat = sardine_truth$structn,
                                 time = fish_times,
                                 species = species,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
catch_structnss_samp <- sample_fish(catch_aggstructnss, effN, sample = FALSE)

catch_resnss_samp <-  sample_fish(catch_aggresnss, effN, sample = FALSE)

# these fishery lengths and weight at age are each output timestep
catch_lengthwt_samp <- calc_age2length(structn = catch_structnss_samp,
                                 resn = catch_resnss_samp,
                                 nums = catch_numsss_samp,
                                 biolprm = sardine_truth$biolprm, 
                                 fgs = sardine_truth$fgs,
                                 maxbin = maxbin,
                                 CVlenage = CVs$lenage, remove.zeroes=TRUE)


```

